---
title: "CogSci Eye-tracking Workshop 2019 - Day 5"
author: "Fabio Trecca"
date: "12/2/2019"
output:
  html_document:
#    theme: paper
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
```

```{r libraries}
library(tidyverse)
library(lme4)
```

## Load data

```{r load data}
## Set working directory to the relevant one on your computer
getwd()
setwd("C:/Users/Bruger/Desktop/Cognitive Science/R_files/Eye-Tracking/Day5")

Fixations <- read_csv('C:/Users/Bruger/Desktop/Cognitive Science/R_files/Eye-Tracking/Day5/fixations2018_2019.csv')
Saccades <- read_csv('C:/Users/Bruger/Desktop/Cognitive Science/R_files/Eye-Tracking/Day5/saccades2018_2019.csv')

# Make sure variables are of the right type
Fixations <- Fixations %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )

Saccades <- Saccades %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Direction = as.character(Direction), #Changed from as.factor to as.character
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )
```
Visual Search:
- VS patterns are affected by task structure and goals (top-down)
- VS patterns are affected by stimulus structure (bottom-up)
- Search direction is affected by cultural constraints (e.g., reading
direction)

. Visual search:
Search patterns are affected by task structure and goals (top-
down)
. Fixations: duration
. Saccades: N, amplitude
. Bonus:
. Searching directions: In Count task, early fixations
haveMeanPosition < median(Position)
. P(next jump == long jump)
```{r}
VSfix = filter(Fixations, Task == "VisualSearch_Count" | Task == "VisualSearch_Star", Fixation != 1, Duration < 7000)
VSsacc = filter(Saccades, Task == "VisualSearch_Count" | Task == "VisualSearch_Star", Saccade != 1)

#Note that we are removing the first fixatoin and the first saccade in each dataset, as there was no fixation cross

VSfix$StartTime = scale(VSfix$StartTime)
VSsacc$StartTime = scale(VSsacc$StartTime)

FixDmodel = glmer(Duration ~ Task + StartTime + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSfix, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

FixDnullmodel1 = glmer(Duration ~ Task + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSfix, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

FixDnullmodel2 = glmer(Duration ~ StartTime + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSfix, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

FixDnoitem = glmer(Duration ~ Task + StartTime + (1|ParticipantID), family = gaussian(link=log), data = VSfix, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))


FixAnova = anova(FixDmodel, FixDnullmodel1)
FixAnova2 = anova(FixDmodel, FixDnullmodel2) #Begge vores fixed effects er significante fra vores nul-modle

FixAnova3 = anova(FixDmodel, FixDnoitem) #Item unders?ger bottom up effekten



summary(FixDmodel)

#Our interaction models were never significant between Task*StartTime

SaccAmodel = glmer(Amplitude ~ Task + StartTime + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

SaccAnullmodel1 = glmer(Amplitude ~ Task +  (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

SaccAnullmodel2 = glmer(Amplitude ~ StartTime + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

Saccnoitem = glmer(Amplitude ~ Task + StartTime + (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

summary(SaccAmodel)

SaccAnova = anova(SaccAmodel, SaccAnullmodel1)
SaccAnova2 = anova(SaccAmodel, SaccAnullmodel2)
#Both models seem good. The null models tell us that each of our fixed effects are significantly different from the null models. 

SaccAnova3 = anova(SaccAmodel, Saccnoitem)
SaccAnova3 #Item makes a difference


#Now Velocity
SaccVmodel = glmer(MeanVelocity ~ Task + StartTime + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

SaccVnullmodel1 = glmer(MeanVelocity ~ Task + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

SaccVnullmodel2 = glmer(MeanVelocity ~ StartTime + (1|Item) +  (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

SaccVnoitem = glmer(MeanVelocity ~ Task + StartTime +  (1|ParticipantID), family = gaussian(link=log), data = VSsacc, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

SaccVAnova = anova(SaccVmodel, SaccVnullmodel1)
SaccVAnova2 = anova(SaccVmodel, SaccVnullmodel2)
SaccVAnova3 = anova(SaccVmodel, SaccVnoitem) #Checking the effect of item

summary(SaccVmodel)

#Start time falls in both saccade models, indicating that the saccades are longer and faster in the beginning of each item. 

```
Inspecting the data
```{r}
ggplot(VSsacc, aes(x = Amplitude, color = Task)) + geom_density()
ggplot(VSsacc, aes(x = MeanVelocity, color = Task)) + geom_density()
ggplot(VSfix, aes (x = Duration, color = Task)) + geom_density()



quantile(VSsacc$Amplitude, na.rm = TRUE) #75% = 4.69



```

Reading direction hypothesis. More read from left to right, and its determined by culture.
```{r}
filter(VSsacc, Task == "VisualSearch_Count") %>% count(Direction)


```

