---
title: "signe_1_eyetracking"
author: "Signe Kløve Kjær"
date: "8/2/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(lme4)
```

VISUAL SEARCH


## Load data

```{r load data}
## Set working directory to the relevant one on your computer
getwd()
setwd('/Users/signeklovekjaer/Documents/CognitiveScience/4.semester/Computational_modeling_for_cognitive_science/Eye_tracking_workshop/Portfolio')

Fixations <- read_csv('fixations2018_2019.csv')
Saccades <- read_csv('saccades2018_2019.csv')

# Make sure variables are of the right type
Fixations <- Fixations %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )

Saccades <- Saccades %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Direction = as.factor(Direction),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )
```

Fixation duration ~ Task + StartTime(if dependent on task, would interact with task) +  (1 + Task | Item) + (1 + Task + StartTime  | Participant)


```{r}
library(lme4)

library(tidyverse)

#fixation duration


#split the dataset 
Fixations_vs <- subset(Fixations, Task == "VisualSearch_Count" | Task == "VisualSearch_Star")

#need to change number of levels
what <- unique(Fixations_vs$Task, incomparables = FALSE)

what

#use factor()
Fixations_vs$Task <- factor(Fixations_vs$Task)

what <- unique(Fixations_vs$Task, incomparables = FALSE)

what #now it works 


Fixations_vs <- mutate(Fixations_vs, sdStartTime = scale(Fixations_vs$StartTime))

model_1 <- glmer(Duration ~ Task + sdStartTime + (1 + Task + sdStartTime | ParticipantID) + (1 + Task | Item), data = Fixations_vs, family = gaussian(link = log))

model_2 <- glmer(Duration ~ Task + (1 + Task + sdStartTime | ParticipantID) + (1 + Task | Item), data = Fixations_vs, family = gaussian(link = log))

sum_model_null <- glmer(Duration ~ 1 + (1 + Task + sdStartTime | ParticipantID) + (1 + Task | Item), data = Fixations_vs, family = gaussian(link = log))

sum_model_1 <- summary(model_1)

sum_model_2 <- summary(model_2)

sum_model_2



```


We need to take the exponential of the estimates in the fixed effects in order to be able to interpret the results. 

Take exponential of the intercept (the mean of the counting task) and add it to the estimates of the fixed effects. Important that you take the whole b + a*x before you take the exponential, so exp(intercept + estimate)



```{r}
#interpret the results
mean_duration_count <- exp(5.69196)
mean_duration_count

change_duration_from_count_to_star <- exp( 5.69196 -0.22903)
change_duration_from_count_to_star

```



To find R-squared use Mumin.

```{r}
library(metafor)
library(lmerTest)

#compare the models with anova 
anova <- anova(sum_model_null, sum_model_1, sum_model_2)


summary(anova)
```


Fixation duration:
-------
Duration ~ Task + StartTime_norm + (1 + Task | Item) + (1 + Task + StartTime_norm | ParticipantID)
Duration ~ Task * StartTime_scaled + (1 + Task + StartTime_scaled | Item) + (1 | ParticipantID) + (0 + Task | ParticipantID)


Saccade amplitude:
-------
Amplitude ~ Task + StartTime_scaled + (1 | Item) + (1 | ParticipantID)
Amplitude ~ Task * StartTime_scaled + (1 + Task | ParticipantID)
Amplitude ~ Task * StartTime_norm + (1 + Task | Item) + (1 + Task | ParticipantID)

Saccade velocity:
-------
SacVel ~ Task + StartTime_scaled + (1 | Item) + (1 | ParticipantID)
SacVel ~ Task * StartTime_scaled + (1 + Task | ParticipantID)


SOCIAL ENGAGEMENT 

```{r}
#merge the dataframes to get the items 

logfile_se <- read_csv("logfilesSE.csv")

logfile_se$ParticipantID <- tolower(logfile_se$ParticipantID)

Fixations_se <- subset(Fixations, Task == "SocialEngagement" & Year == 2018)

logfile_se <- read_csv("logfilesSE.csv")

Fixations_se <- merge(Fixations_se, logfile_se, by = c("ParticipantID", "Trial"))


```


```{r}
#extract from filename and prepare dataset

Fixations_se$Condition_di <- ifelse(grepl("div", Fixations_se$Video), "div", "dir")

Fixations_se$Condition_o <- ifelse(grepl( "-", Fixations_se$Video), "-o", "+o")

Fixations_se$video_gender <- ifelse(grepl("f", Fixations_se$Video), "f", "m")

Fixations_se$Condition <- paste(Fixations_se$Condition_di,Fixations_se$Condition_o, sep = "")

#deselect extra columns 
Fixations_se <- subset(Fixations_se, select = -c(Task.y))

Fixations_se$Task.x <- factor(Fixations_se$Task.x)

#scale starttime
Fixations_se <- mutate(Fixations_se, stdStartTime = scale(Fixations_se$StartTime))


```

```{r}
#what kind of model should we make?

library(ggplot2)
hist(Fixations_se$PupilSize)

ggplot(Fixations_se, aes(PupilSize)) + geom_density() 


ggplot(Fixations_se, aes(PupilSize, color = ParticipantID)) + geom_density()


```

From plotting distribution of pupilsize for each participant it's clear that the plot of all participants pupil distribution is mostly driven by individuals having different baselines. The distributions for the individual participants look somewhat like a gaussian distribution. We will therefore create a lmer model assuming a gaussian distribution and controlling for participants having different baselines by including Participant ID as random intercepts.   

```{r}
#create models

model_pupil <- lmer(PupilSize ~ Condition + stdStartTime + (1 + Condition + stdStartTime| ParticipantID), data = Fixations_se, REML = FALSE)

model_null <-  lmer(PupilSize ~ Condition + (1 + Condition + stdStartTime| ParticipantID), data = Fixations_se, REML = FALSE)

model_pupil_inter <- lmer(PupilSize ~ Condition_o*Condition_di + stdStartTime + (1 + Condition + stdStartTime| ParticipantID), data = Fixations_se, REML = FALSE)



summary(model_pupil)

summary(model_pupil_inter)


model_fix <- glmer(Duration ~ Condition +  stdStartTime + (1 + Condition +  stdStartTime | ParticipantID), family = gaussian (link = log), data = Fixations_se)

summary(model_fix)

library(MuMIn)

r.squaredGLMM(model_pupil)

r.squaredGLMM(model_null)

```
```{r}
#create models wih separate conditions

model_pupil <- glmer(PupilSize ~ Condition_o + Condition_di + stdStartTime + (1 + Condition_o + Condition_di + stdStartTime| ParticipantID),family = gaussian (link = log), data = Fixations_se)


model_pupil_1 <- glmer(PupilSize ~ Condition_o + Condition_di + stdStartTime + (1 + stdStartTime| ParticipantID),family = gaussian (link = log), data = Fixations_se)


model_null <-  glmer(PupilSize ~ Condition_o + Condition_di + (1 + Condition_o + Condition_di + stdStartTime| ParticipantID),family = gaussian (link = log), data = Fixations_se)

model_pupil_inter <- glmer(PupilSize ~ Condition_o*Condition_di + stdStartTime + (1 + Condition_di + Condition_o + stdStartTime| ParticipantID),family = gaussian (link = log), data = Fixations_se)


model_pupil_inter_1 <- glmer(PupilSize ~ Condition_o*Condition_di + stdStartTime + (1 + stdStartTime| ParticipantID),family = gaussian (link = log), data = Fixations_se)

summary(model_pupil_inter_1)

summary(model_pupil_1)



model_fix <- glmer(Duration ~ Condition_di * Condition_o +  stdStartTime + (1 + Condition_di + Condition_o +  stdStartTime | ParticipantID), family = gaussian (link = log), data = Fixations_se)

summary(model_fix)


r.squaredGLMM(model_pupil_1)

r.squaredGLMM(model_pupil)
```







```{r}
#what kind of participants

unique(Fixations_se$ParticipantID)

Fixations_se$video_gender <- as.factor(Fixations_se$video_gender)

unique(Fixations_se$ParticipantID) %>%
  table(Fixations_se$video_gender)

unique(Fixations_se$ParticipantID)

```


```{r}
#plot predictions of pupil size 
library(beeswarm)
library(ggplot2)
library(ggbeeswarm)

pred_pup = Fixations_se %>%
  distinct(Condition_di, Condition_o, stdStartTime, ParticipantID, PupilSize) %>%
  cbind(predict(model_pupil_inter_1, newdata_1 = ., interval = "confidence"))


plot_pupil_pred <- ggplot(Fixations_se, aes(x = Condition_di, y = PupilSize)) + 
  geom_quasirandom(alpha = 0.5, colour = "dark red") + 
  labs(x = "Condition (directed, diverse) ", y = "Pupil size", title = "Mean model predictions of pupil size") + 

  #geom_boxplot(alpha = 0.5) +p
  #geom_point(aes(y = `predict(model_pupil, newdata_1 = ., interval = "confidence")`), data = pred_pup, size = 2) +
  stat_summary(fun.data = "mean_cl_boot", colour = "red", size = 1) + 
  #geom_errorbar(aes(y = NULL, ymin = lwr, ymax = upr), data = pred_RR_HR, width = .3) +
  facet_wrap(~Condition_o) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_pupil_pred




```


FIXATION DURATION 

```{r}


#predictions 
pred_fix = Fixations_se %>%
  distinct(Condition_o, Condition_di, stdStartTime, ParticipantID, Duration) %>%
  cbind(predict(model_fix, newdata = ., interval = "confidence"))


plot_pred_fix <- ggplot(Fixations_se, aes(x = Condition_di, y = Duration)) + 
  geom_quasirandom(alpha = 0.5, colour = "dark red") + 
  labs(x = "Condition (directed, diverse)", y = "Fixation duration", title = "Mean model predictions og fixation duration") + 

  #geom_boxplot(alpha = 0.5) +p
  #geom_point(aes(y = `predict(model_fix, newdata = ., interval = "confidence")`), data = pred_fix, size = 2) 
  stat_summary(fun.data = "mean_cl_boot", colour = "red", size = 0.5) + 
  #geom_errorbar(aes(y = NULL, ymin = lwr, ymax = upr), data = pred_RR_HR, width = .3) +
  facet_wrap(~Condition_o) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_pred_fix




```
