---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate (more than one predictor) linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS

```{r}
library(rethinking)

#load data
asd <- read.csv("Assignment3Data.csv")

#only use visit 1
asd_2 <- asd[ asd$Visit <= 1 , ]

#only use children with asd
asd_2 <- asd_2[ asd_2$ASD >= 1 , ]


#remove child with ID 1, because it contain NA values
asd_2 <- asd_2[ asd_2$ChildID >= 2 , ]


#lets center the predictor variable to avoid correlation between alpha and beta 

asd_2$NonVerbalIQ_c <- scale(asd_2$NonVerbalIQ)
asd_2$VerbalIQ_c <- scale(asd_2$VerbalIQ)
asd_2$SocialIQ_c <- scale(asd_2$SocialIQ)


mean(asd_2$VerbalIQ_c)
mean(asd_2$NonVerbalIQ_c)
mean(asd_2$SocialIQ_c)
```



```{r}
#plot of data
dens(asd_2$ADOS) #for each value of verbalIQ/NonVerbalIQ/SocialIQ there will be a normal distribution, menaing that the residuals will be normal distributed 

```

Priors:

Autism cannot be negative 

Make prior predictive plot --> run the model without the likelyhood (data) --> called a prior predictive check 


```{r}
#make prior predictive check --> run the model without the likelyhood (data)

library(brms)

#Define the model
m_verbal <- bf(ADOS ~ 1 + VerbalIQ_c)

# Ask which priors need to be defined, but I'm not going to use it  
get_prior(m_verbal, data = asd_2, family = gaussian)

# Define the prior
#intecept is the ados you have if you are average in verbal iq
prior = c(prior(normal(12,5), class = "Intercept"), prior(normal(-5,5), class = "b") , prior(normal(0,5), class = "sigma"))

# Draw the consequences of priors 
m_verbal_prior <- brm(m_verbal, asd_2, prior = prior, sample_prior = "only")
# Prior Predictive Check
pp_check(m_verbal_prior)
labels()
# Repeat until you are satisfied

```

```{r}
#VERBAL_IQ

library(rethinking)


m1.1 <- rethinking::map(
    alist(
        ADOS ~ dnorm( mu , sigma ) ,
        mu <- a + b*VerbalIQ_c ,
        a ~ dnorm( 12 , 5 ) ,
        b ~ dnorm( -5 , 5 ) ,
        sigma ~ dunif( 0 , 5 ) ),
data = asd_2 )

precis(m1.1)


#look at the fit maximum of the model
m1.1_estimates <- precis(m1.1)


```

```{r}
#plot model of verbal_iq

# define sequence of verbaliq to compute predictions for
# these values will be on the horizontal axis
verbal.seq <- seq( from=-3 , to=3 , by=1 )

#summarise distribution of mu

mu <- link (m1.1, data = data.frame(VerbalIQ_c = verbal.seq))


# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

#simulate ADOS
sim.ADOS <- sim(m1.1, data = list(VerbalIQ_c = verbal.seq))
ADOS.PI <- apply(sim.ADOS, 2, PI, prob = 0.89)

# plot raw data
plot( ADOS ~ VerbalIQ_c , asd_2 , col=col.alpha(rangi2,0.5) )
# draw MAP line
lines( verbal.seq , mu.mean )
# draw HPDI region for line
shade( mu.HPDI , verbal.seq )
# draw PI region for simulated heights
shade( ADOS.PI , verbal.seq )
```



```{r}
#NONVERBAL_IQ

m1.2 <- rethinking::map(
    alist(
        ADOS ~ dnorm( mu , sigma ) ,
        mu <- a + b*NonVerbalIQ_c ,
        a ~ dnorm( 12 , 5 ) ,
        b ~ dnorm( -5 , 5 ) ,
        sigma ~ dnorm( 0 , 5 )
), data = asd_2 )



#look at the fit maximum of the model
m1.2_estimates <- precis(m1.2)
```

```{r}
#plot model of nonverbal_iq

# define sequence of nonverbaliq to compute predictions for
# these values will be on the horizontal axis
nonverbal.seq <- seq( from=-3 , to=3 , by=1 )

#summarise distribution of mu

mu <- link (m1.2, data = data.frame(NonVerbalIQ_c = nonverbal.seq))


# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

#simulate ADOS
sim.ADOS <- sim(m1.2, data = list(NonVerbalIQ_c = nonverbal.seq))
ADOS.PI <- apply(sim.ADOS, 2, PI, prob = 0.89)

# plot raw data
plot( ADOS ~ NonVerbalIQ_c , asd_2 , col=col.alpha(rangi2,0.5) )
# draw MAP line
lines( nonverbal.seq , mu.mean )
# draw HPDI region for line
shade( mu.HPDI , nonverbal.seq )
# draw PI region for simulated heights
shade( ADOS.PI , nonverbal.seq )
```




```{r}
#SOCIAL_IQ

m1.3 <- rethinking::map(
    alist(
        ADOS ~ dnorm( mu , sigma ) ,
        mu <- a + b*SocialIQ_c ,
        a ~ dnorm( 12 , 5 ) ,
        b ~ dnorm( -5 , 5 ) ,
sigma ~ dunif( 0 , 5 ) ),
data = asd_2)




#look at the fit maximum of the model
m1.3_estimates <- precis(m1.3)
```


```{r}
#social_IQ


# define sequence of nonverbaliq to compute predictions for
# these values will be on the horizontal axis
social.seq <- seq( from=-3 , to=3 , by=1 )

#summarise distribution of mu

mu <- link (m1.3, data = data.frame(SocialIQ_c = social.seq))


# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

#simulate ADOS
sim.ADOS <- sim(m1.3, data = list(SocialIQ_c = social.seq))
ADOS.PI <- apply(sim.ADOS, 2, PI, prob = 0.89)

# plot raw data
plot( ADOS ~ SocialIQ_c , asd_2 , col=col.alpha(rangi2,0.5) )
# draw MAP line
lines( social.seq , mu.mean )
# draw HPDI region for line
shade( mu.HPDI , social.seq )
# draw PI region for simulated heights
shade( ADOS.PI , social.seq )






```


ROBUSTNESS CHECK

We tried with different values for mean alpha and mean beta. 

```{r}

#define the model
flist_prior_check_1.1 <- alist(
  ADOS ~ dnorm( mu , sigma ) ,
        mu <- a + b*SocialIQ_c ,
        a ~ dnorm( 5 , 5 ) ,
        b ~ dnorm(-20, 5 ) ,
        sigma ~ dnorm( 0 , 5 )
)


#fit the model to the data
check_1.1 <- map(flist_prior_check_1.1, data = asd_2)


#look at the fit maximum of the model
precis(check_1.1)


```




2. Do the different aspects of IQ account for different portions of the variance in ADOS? 


2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?

It could make sense if they do not correlate too much. Use chapter 5 about multivariate models to assess whether they correlate. Non-verbal_IQ and social_IQ could potentially explain different variance in the data. 

The first thing we will do is to check whether the IQ measures correlate. 

```{r}
#Create model with all predictors


m1.4 <- rethinking::map(
    alist(
        ADOS ~ dnorm( mu , sigma ) ,
        mu <- a + bs*SocialIQ_c + bn*NonVerbalIQ_c + bv*VerbalIQ_c ,
        a ~ dnorm( 8 , 5 ) ,
        bs ~ dnorm(0, 5 ) ,
        bn ~ dnorm(0,5),
        bv ~ dnorm(0,5),
        sigma ~ dnorm( 0 , 5 )
), data = asd_2)



pairs(~  ADOS + VerbalIQ_c + NonVerbalIQ_c + SocialIQ_c, 
      data = asd_2, col = rangi2)
```




```{r}
cor(asd_2$VerbalIQ_c, asd_2$NonVerbalIQ_c)

cor(asd_2$VerbalIQ_c, asd_2$SocialIQ_c)

cor(asd_2$NonVerbalIQ_c, asd_2$SocialIQ_c)


```

After checking the pairs plot and the estimated correlation between the predictor variable, we can see that verbal and nonverbal iq is correlated by 0.68 and with social iq by 0.61. This we will keep in mind as we procede keeping all the variables as predictors in the model. 

```{r}
precis(m1.4)
```



2.2. Build the model, assess its quality, write a few lines interpreting the results.

Look at table of estimates and posterior distribution. 

Assess it's quality: Create predictive posteiror plot and look at error or make shaded areas around model and data 

Now we want to plot the predictor residual plots


```{r}
#compute the residuals

#we have standarddized the IQ measures, so we expect alpha to be around 0 

#create model investigating verbal 
#define the model

m1.5 <- rethinking::map(
    alist(
        VerbalIQ_c ~ dnorm( mu , sigma ) ,
        mu <- a + bs*SocialIQ_c + bn*NonVerbalIQ_c,
        a ~ dnorm( 0, 5 ) ,
        bs ~ dnorm(0, 5 ) ,
        bn ~ dnorm(0,5),
        sigma ~ dnorm( 0 , 5 )
), data = asd_2)




# compute expected value at MAP, for each State
mu_verbal <- coef(m1.5)['a'] + coef(m1.5)['bs']*asd_2$SocialIQ_c + coef(m1.5)['bn']*asd_2$NonVerbalIQ_c
# compute residual for each State
m.resid_verbal <- asd_2$VerbalIQ_c - mu_verbal



```

```{r}
#plot the residual
plot(x = m.resid_verbal, y = asd_2$ADOS)
```



```{r}
#compute the residuals

#create model investigating nonverbal 
#define the model

m1.6 <- rethinking::map(
    alist(
        NonVerbalIQ_c ~ dnorm( mu , sigma ) ,
        mu <- a + bs*SocialIQ_c + bv*VerbalIQ_c,
        a ~ dnorm( 0 , 5 ) ,
        bs ~ dnorm(0, 5 ) ,
        bv ~ dnorm(0,5),
        sigma ~ dnorm( 0 , 5 )
), data = asd_2)


# compute expected value at MAP, for each State
mu_nonverbal <- coef(m1.6)['a'] + coef(m1.6)['bs']*asd_2$SocialIQ_c + coef(m1.6)['bv']*asd_2$VerbalIQ_c
# compute residual for each State
m.resid_nonverbal <- asd_2$NonVerbalIQ_c - mu_nonverbal



```


```{r}
#plot the residual
plot(x = m.resid_nonverbal, y = asd_2$ADOS)
```

```{r}
#compute the residuals

#create model investigating social IQ 
#define the model

m1.7 <- rethinking::map(
    alist(
        SocialIQ_c ~ dnorm( mu , sigma ) ,
        mu <- a + bn*NonVerbalIQ_c + bv*VerbalIQ_c,
        a ~ dnorm( 0 , 5 ) ,
        bn ~ dnorm(0, 5 ) ,
        bv ~ dnorm(0,5),
        sigma ~ dnorm( 0 , 5 )
), data = asd_2)


# compute expected value at MAP, for each State
mu_social <- coef(m1.7)['a'] + coef(m1.7)['bn']*asd_2$NonVerbalIQ_c + coef(m1.7)['bv']*asd_2$VerbalIQ_c
# compute residual for each State
m.resid_social <- asd_2$NonVerbalIQ_c - mu_social



```


```{r}
#plot the residual
plot(x = m.resid_social, y = asd_2$ADOS)
```




```{r}
#COunterfactual plot showing the impact of changes in Social IQ 


# prepare new counterfactual data
A.avg <- mean( d$MedianAgeMarriage.s )
R.seq <- seq( from=-3 , to=3 , length.out=30 )
pred.data <- data.frame(
    Marriage.s=R.seq,
    MedianAgeMarriage.s=A.avg
)
# compute counterfactual mean divorce (mu)
mu <- link( m5.3 , data=pred.data )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI )
# simulate counterfactual divorce outcomes
R.sim <- sim( m5.3 , data=pred.data , n=1e4 )

R.PI <- apply( R.sim , 2 , PI )
# display predictions, hiding raw data with type="n"
plot( Divorce ~ Marriage.s , data=d , type="n" )
mtext( "MedianAgeMarriage.s = 0" )
lines( R.seq , mu.mean )
shade( mu.PI , R.seq )
shade( R.PI , R.seq )

```





3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

```{r}
#only use visit 1
asd_3 <- asd[ asd$Visit <= 1 , ]

table(asd_3$ASD)

table(asd_3$ADOS)
```

It does not really make sense as the distribtion of ADOS will have a bump around 0, as 25 of 38 of the TD children have an ADOS score of 0.Thereby mapping ADOS as a continious relation is a little counter intuitive, as there actually is two categories hidden inside a continious variable. 

However autism is a spectrum disorder and in that case it makes sense to have ADOS as a continious variable. 

So we will try and build the models and see how it works. 


```{r}
#lets center the predictor variable to avoid correlation between alpha and beta 

asd_3$NonVerbalIQ_c <- scale(asd_3$NonVerbalIQ)
asd_3$VerbalIQ_c <- scale(asd_3$VerbalIQ)
asd_3$SocialIQ_c <- scale(asd_3$SocialIQ)

#remove child with ID 1, because it contain NA values
asd_3 <- asd_3[ asd_3$ChildID >= 2 , ]


#make prior predictive check --> run the model without the likelyhood (data)

library(brms)

#Define the model
m_full <- bf(ADOS ~ 1 + VerbalIQ_c + NonVerbalIQ_c + SocialIQ_c)


# Define the prior
#intecept is the ados you have if you are average in verbal iq
prior = c(prior(normal(6,5), class = "Intercept"), prior(normal(0,5), class = "b"), prior(normal(0,5), class = "sigma"))

# Draw the consequences of priors 
m_full_prior <- brm(m_full, asd_3, prior = prior, sample_prior = "only")
# Prior Predictive Check
pp_check(m_full_prior)
# Repeat until you are satisfied

```

```{r}
m_full_1 <- rethinking::map(
    alist(
        ADOS ~ dnorm( mu , sigma ) ,
        mu <- a + bv*VerbalIQ_c + bn*NonVerbalIQ_c + bs*SocialIQ_c ,
        a ~ dnorm( 6 , 5 ),
        bn ~ dnorm( 0 , 5 ),
        bs ~ dnorm( 0 , 5 ), 
        bv ~ dnorm( 0 , 5 ),
        sigma ~ dnorm( 0 , 5 )
), data=asd_3)


precis(m_full_1)
```

The beta values for nonverbal and verbal crosses zero. The social iq looks more reliable as the confidence intervals are on the same side of zero and is not too far spread out. 

Lets try and make a counterfactual plot

```{r}
#COunterfactual plot showing the impact of changes in Marriage.s

# prepare new counterfactual data
nonverbal.avg <- mean( asd_3$NonVerbalIQ_c)
verbal.avg <- mean( asd_3$VerbalIQ_c)
Social.seq <- seq( from=-3 , to=3 , length.out=30 )
pred.data <- data.frame(
    SocialIQ_c=Social.seq,
    NonVerbalIQ_c=nonverbal.avg,
    VerbalIQ_c = verbal.avg
)

# compute counterfactual mean ados (mu)
mu <- link( m_full_1 , data=asd_3 )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI )
# simulate counterfactual divorce outcomes
R.sim <- sim( m_full_1 , data=pred.data , n=1e4 )

R.PI <- apply( R.sim , 2 , PI )
# display predictions, hiding raw data with type="n"
plot( ADOS ~ SocialIQ_c , data=asd_3 , type="n" )
mtext( "Nonverbal = 0, Verbal = 0" )
lines( Social.seq , mu.mean )
shade( mu.PI , Social.seq )
shade( R.PI , Social.seq )


Social.seq
```



Would maybe be better to create a model with asd predicted by verbal_iq. 



4. Let's discuss contents:



4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?

Social IQ explains some variance which is not captured by nonverbal and verbal IQ. 



4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.




5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


