---
title: "Assignment3"
author: "Simon Moeller Nielsen"
date: "5/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS
```{r 1. setup}
library(rstan)
library(tidyverse)
library(rethinking)
library(brms)
setwd('C:/Users/slmoni/Documents/Uni/Stats/ComMod/3_multiple_reg_and_causal_inferences')
asd<-read.csv('Assignment3Data.csv')

# why scaling? To avoid correlation between a and b, as well as making comperision
asd1<-filter(asd, Visit==1)
asd1<-filter(asd1, ChildID!=1)
asd1<-filter(asd1, ASD==1)
#asd1<-filter(asd1, ADOS>1) #removing child with 0 ADOS since he is an outlier

asd2<-filter(asd, Visit==1)
asd2<-filter(asd2, ChildID!=1)

asd3<-filter(asd, Visit==1)
asd3<-filter(asd3, ChildID!=1)
asd3<-filter(asd3, ASD==0)

asd1$NonVerbalIQ.s<-(asd1$NonVerbalIQ-mean(asd1$NonVerbalIQ,na.rm=T))/sd(asd1$NonVerbalIQ,na.rm=T)
asd1$VerbalIQ.s<-(asd1$VerbalIQ-mean(asd1$VerbalIQ,na.rm=T))/sd(asd1$VerbalIQ,na.rm=T)
asd1$SocialIQ.s<-(asd1$SocialIQ-mean(asd1$SocialIQ,na.rm=T))/sd(asd1$SocialIQ,na.rm=T)
#asd1$ADOS.s<-(asd1$ADOS-mean(asd1$ADOS,na.rm=T))/sd(asd1$ADOS,na.rm=T)

asd2$NonVerbalIQ.s<-(asd2$NonVerbalIQ-mean(asd2$NonVerbalIQ,na.rm=T))/sd(asd2$NonVerbalIQ,na.rm=T)
asd2$VerbalIQ.s<-(asd2$VerbalIQ-mean(asd2$VerbalIQ,na.rm=T))/sd(asd2$VerbalIQ,na.rm=T)
asd2$SocialIQ.s<-(asd2$SocialIQ-mean(asd2$SocialIQ,na.rm=T))/sd(asd2$SocialIQ,na.rm=T)
#asd2$ADOS.s<-log(asd2$ADOS)
  
asd3$NonVerbalIQ.s<-(asd3$NonVerbalIQ-mean(asd3$NonVerbalIQ,na.rm=T))/sd(asd3$NonVerbalIQ,na.rm=T)
asd3$VerbalIQ.s<-(asd3$VerbalIQ-mean(asd3$VerbalIQ,na.rm=T))/sd(asd3$VerbalIQ,na.rm=T)
asd3$SocialIQ.s<-(asd3$SocialIQ-mean(asd3$SocialIQ,na.rm=T))/sd(asd3$SocialIQ,na.rm=T)

asd1$Estimate<- asd1$ADOS
asd2$Estimate<- asd2$ADOS
```

```{r 1.1}
## VerbalIQ model
verb <- bf(ADOS ~ 1 + VerbalIQ.s)

prior <- c(prior(normal( 12 , 5 ), class='Intercept'), prior(normal( 0 , 5 ), class='b'), prior(normal( 0 , 5 ), class='sigma'))
post_verb <- brm(verb, asd1, prior = prior, sample_prior = T)

prior_verb <- brm(verb, asd1, prior = prior, sample_prior = 'only')
pp_check(prior_verb)

summary(post_verb)

# Plot
# We need new `nd` data
nd <- 
  tibble(VerbalIQ.s          = seq(from = -3, to = 3, length.out = 30))#,
         #NonVerbalIQ.s = rep(mean(asd1$NonVerbalIQ.s), times = 30))
  
fitted(post_verb, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_verb, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = VerbalIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd1$VerbalIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nNonVerbalIQ.s = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd1) 

fitted(post_verb) %>%
  as_tibble() %>%
  bind_cols(asd1) %>%
  
  ggplot(aes(x = ADOS, y = Estimate)) +
  geom_abline(linetype = 2, color = "grey50", size = .5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 3/4) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5),
                 size = 1/4, color = "firebrick4") +
  geom_linerange(aes(ymin = Estimate - Est.Error, 
                     ymax = Estimate + Est.Error),
                 size = 1/2, color = "firebrick4") +
  # Note our use of the dot placeholder, here: https://magrittr.tidyverse.org/reference/pipe.html
  # geom_text(data = . %>% filter(Loc %in% c("ID", "UT")),
            # aes(label = Loc), 
            #hjust = 0, nudge_x = - 0.65) +
  labs(title='Posterior predictive plot for VerbalIQ',x = "Observed ADOS", y = "Predicted ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())


```

```{r 1.2}
## NonVerbalIQ model 
nonverb <- bf(ADOS ~ 1 + NonVerbalIQ.s)

prior <- c(prior(normal( 12 , 5 ), class='Intercept'), prior(normal( 0 , 5 ), class='b'), prior(normal( 0 , 5 ), class='sigma'))
post_nonverb <- brm(nonverb, asd1, prior = prior, sample_prior = T)

pp_check(post_nonverb)

summary(post_nonverb)

# Plot
# We need new `nd` data
nd <- 
  tibble(NonVerbalIQ.s          = seq(from = -3, to = 3, length.out = 30))#,
         #NonVerbalIQ.s = rep(mean(asd1$NonVerbalIQ.s), times = 30))
  
fitted(post_nonverb, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_nonverb, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = NonVerbalIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd1$NonVerbalIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nNonVerbalIQ.s = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd1) 

fitted(post_nonverb) %>%
  as_tibble() %>%
  bind_cols(asd1) %>%
  
  ggplot(aes(x = ADOS, y = Estimate)) +
  geom_abline(linetype = 2, color = "grey50", size = .5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 3/4) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5),
                 size = 1/4, color = "firebrick4") +
  geom_linerange(aes(ymin = Estimate - Est.Error, 
                     ymax = Estimate + Est.Error),
                 size = 1/2, color = "firebrick4") +
  # Note our use of the dot placeholder, here: https://magrittr.tidyverse.org/reference/pipe.html
  # geom_text(data = . %>% filter(Loc %in% c("ID", "UT")),
            # aes(label = Loc), 
            #hjust = 0, nudge_x = - 0.65) +
  labs(title='Posterior predictive plot for NonverbalIQ',x = "Observed ADOS", y = "Predicted ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

```{r 1.3}
## SocialIQ model 
social <- bf(ADOS ~ 1 + SocialIQ.s)

prior <- c(prior(normal( 12 , 5 ), class='Intercept'), prior(normal( 0 , 5 ), class='b'), prior(normal( 0 , 5 ), class='sigma'))
post_social <- brm(social, asd1, prior = prior, sample_prior = T)

pp_check(post_social)

summary(post_social)

# Plot
# We need new `nd` data
nd <- 
  tibble(SocialIQ.s          = seq(from = -3, to = 3, length.out = 30))#,
         #NonVerbalIQ.s = rep(mean(asd1$NonVerbalIQ.s), times = 30))
  
fitted(post_social, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_social, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = SocialIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd1$SocialIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nNonVerbalIQ.s = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd1) 

fitted(post_social) %>%
  as_tibble() %>%
  bind_cols(asd1) %>%
  
  ggplot(aes(x = ADOS, y = Estimate)) +
  geom_abline(linetype = 2, color = "grey50", size = .5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 3/4) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5),
                 size = 1/4, color = "firebrick4") +
  geom_linerange(aes(ymin = Estimate - Est.Error, 
                     ymax = Estimate + Est.Error),
                 size = 1/2, color = "firebrick4") +
  # Note our use of the dot placeholder, here: https://magrittr.tidyverse.org/reference/pipe.html
  # geom_text(data = . %>% filter(Loc %in% c("ID", "UT")),
            # aes(label = Loc), 
            #hjust = 0, nudge_x = - 0.65) +
  labs(title='Posterior predictive plot for SocialIQ', x = "Observed ADOS", y = "Predicted ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())

```

2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
```{r 2.}
# Correlation tests
cor(asd1$NonVerbalIQ.s,asd1$VerbalIQ.s)
cor(asd1$NonVerbalIQ.s,asd1$SocialIQ.s)
cor(asd1$VerbalIQ.s,asd1$SocialIQ.s)

pairlist()
```
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.

```{r 2.}
## All model 
all <- bf(ADOS ~ 1 + SocialIQ.s+NonVerbalIQ.s+VerbalIQ.s)

prior <- c(prior(normal( 12 , 5 ), class='Intercept'), prior(normal( -2 , 5 ), class='b'), prior(normal( 0 , 2 ), class='sigma'))
post_all <- brm(all, asd1, prior = prior, sample_prior = T)

pp_check(post_all)

summary(post_all)
```

```{r 2.plot SocialIQ}
# Plot
# We need new `nd` data
nd <- tibble( SocialIQ.s      = seq(from = -3, to = 3, length.out = 30),
         NonVerbalIQ.s = rep(mean(asd1$NonVerbalIQ.s), times = 30),
         VerbalIQ.s = rep(mean(asd1$VerbalIQ.s), times = 30))
  
fitted(post_all, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_all, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = SocialIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd1$SocialIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nNonVerbalIQ = 0 & VerbalIQ = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd1) 

```
```{r 2. plot VerbalIQ}
nd <- tibble(  VerbalIQ.s     = seq(from = -3, to = 3, length.out = 30),
         NonVerbalIQ.s = rep(mean(asd1$NonVerbalIQ.s), times = 30),
         SocialIQ.s = rep(mean(asd1$SocialIQ.s), times = 30))
  
fitted(post_all, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_all, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = VerbalIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd1$SocialIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nNonVerbalIQ = 0 & SocialIQ = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd1) 
```
```{r 2. plot NonVerbalIQ}
nd <- tibble( NonVerbalIQ.s   = seq(from = -3, to = 3, length.out = 30),
         VerbalIQ.s = rep(mean(asd1$VerbalIQ.s), times = 30),
         SocialIQ.s = rep(mean(asd1$SocialIQ.s), times = 30))
  
fitted(post_all, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_all, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = NonVerbalIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd1$SocialIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nVerbalIQ = 0 & SocialIQ = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd1) 
```
```{r 2. posterior predictive plot}
fitted(post_all) %>%
  as_tibble() %>%
  bind_cols(asd1) %>%
  
  ggplot(aes(x = ADOS, y = Estimate)) +
  geom_abline(linetype = 2, color = "grey50", size = .5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 3/4) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5),
                 size = 1/4, color = "firebrick4") +
  geom_linerange(aes(ymin = Estimate - Est.Error, 
                     ymax = Estimate + Est.Error),
                 size = 1/2, color = "firebrick4") +
  # Note our use of the dot placeholder, here: https://magrittr.tidyverse.org/reference/pipe.html
  # geom_text(data = . %>% filter(Loc %in% c("ID", "UT")),
            # aes(label = Loc), 
            #hjust = 0, nudge_x = - 0.65) +
  labs(title='Posterior predictive plot for the model with all predictors',x = "Observed ADOS", y = "Predicted ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())
```
3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.
```{r 3}
## All model 
all <- bf(ADOS ~ 1 + SocialIQ.s+NonVerbalIQ.s+VerbalIQ.s)

prior <- c(prior(normal( 12 , 5 ), class='Intercept'), prior(normal( 0 , 5 ), class='b'), prior(normal( 0 , 5 ), class='sigma'))

prior_fullall <- brm(all, asd2, prior = prior, sample_prior = 'only')

pp_check(prior_fullall)

post_fullall <- brm(all, asd2, prior = prior, sample_prior = T)

summary(post_fullall)
```

```{r 3.plot SocialIQ}
# Plot
# We need new `nd` data
nd <- tibble( SocialIQ.s      = seq(from = -3, to = 3, length.out = 30),
         NonVerbalIQ.s = rep(mean(asd2$NonVerbalIQ.s), times = 30),
         VerbalIQ.s = rep(mean(asd2$VerbalIQ.s), times = 30))
  
fitted(post_fullall, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_fullall, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = SocialIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd2$SocialIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nNonVerbalIQ = 0 & VerbalIQ = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd2) 

```
```{r 3. plot VerbalIQ}
nd <- tibble(  VerbalIQ.s     = seq(from = -3, to = 3, length.out = 30),
         NonVerbalIQ.s = rep(mean(asd2$NonVerbalIQ.s), times = 30),
         SocialIQ.s = rep(mean(asd2$SocialIQ.s), times = 30))
  
fitted(post_fullall, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_fullall, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = VerbalIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd2$VerbalIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nNonVerbalIQ = 0 & SocialIQ = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd2) 
```
```{r 3. plot NonVerbalIQ}
nd <- tibble( NonVerbalIQ.s   = seq(from = -3, to = 3, length.out = 30),
         VerbalIQ.s = rep(mean(asd2$VerbalIQ.s), times = 30),
         SocialIQ.s = rep(mean(asd2$SocialIQ.s), times = 30))
  
fitted(post_fullall, newdata = nd) %>% 
  as_tibble() %>% 
  # Since `fitted()` and `predict()` name their intervals the same way, we'll need to
  # `rename()` then to keep them straight.
  rename(f_ll = Q2.5,
         f_ul = Q97.5) %>% 
  # Note how we're just nesting the `predict()` code right inside `bind_cols()`
  bind_cols(
    predict(post_fullall, newdata = nd) %>% 
      as_tibble() %>% 
      # Since we only need the intervals, we'll use `transmute()` rather than `mutate()`
      transmute(p_ll = Q2.5,
                p_ul = Q97.5)
  ) %>% 
  bind_cols(nd) %>% 
  
  # We're finally ready to plot
  ggplot(aes(x = NonVerbalIQ.s, y = Estimate)) +
  geom_ribbon(aes(ymin = p_ll, ymax = p_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_ribbon(aes(ymin = f_ll, ymax = f_ul),
              fill = "firebrick", alpha = 1/5) +
  geom_line(color = "firebrick4") +
  coord_cartesian(xlim = range(asd2$NonVerbalIQ.s)) +
  labs(subtitle = "Counterfactual plot for which\nVerbalIQ = 0 & SocialIQ = 0",
       y = "ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())+
  geom_point(data = asd2) 
```
```{r 3. posterior predictive plot}
fitted(post_fullall) %>%
  as_tibble() %>%
  bind_cols(asd2) %>%
  
  ggplot(aes(x = ADOS, y = Estimate)) +
  geom_abline(linetype = 2, color = "grey50", size = .5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 3/4) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5),
                 size = 1/4, color = "firebrick4") +
  geom_linerange(aes(ymin = Estimate - Est.Error, 
                     ymax = Estimate + Est.Error),
                 size = 1/2, color = "firebrick4") +
  # Note our use of the dot placeholder, here: https://magrittr.tidyverse.org/reference/pipe.html
  # geom_text(data = . %>% filter(Loc %in% c("ID", "UT")),
            # aes(label = Loc), 
            #hjust = 0, nudge_x = - 0.65) +
  labs(title='Posterior predictive plot for the model with all predictors',x = "Observed ADOS", y = "Predicted ADOS") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

```{r 3}
## All model 
all <- bf(ADOS ~ 1 + SocialIQ.s+NonVerbalIQ.s+VerbalIQ.s)

prior <- c(prior(normal( 12 , 5 ), class='Intercept'), prior(normal( 0 , 5 ), class='b'), prior(normal( 0 , 5 ), class='sigma'))

prior_fullall <- brm(all, asd2, prior = prior, sample_prior = 'only')

pp_check(prior_fullall)

post_fullall <- brm(all, asd2, prior = prior, sample_prior = T)

summary(post_fullall)
```

```{r 4}
cor(asd1$NonVerbalIQ.s,asd1$VerbalIQ.s)
cor(asd1$NonVerbalIQ.s,asd1$SocialIQ.s)
cor(asd1$VerbalIQ.s,asd1$SocialIQ.s)

cor(asd3$NonVerbalIQ.s,asd3$VerbalIQ.s)
cor(asd3$NonVerbalIQ.s,asd3$SocialIQ.s)
cor(asd3$VerbalIQ.s,asd3$SocialIQ.s)

```


4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


