---
title: "hand_priming_exercises"
author: "Signe Kløve Kjær"
date: "3/4/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. 

Hypothesis: 

- In the condition with high targets and one low target, we expect the movement above the low target to be higher than the movement above the condition with only low targerts. 

- In the condition with low targets and one high targets, we expect the lower targets to be overshot relative to the condition with low targets --> preparing to move above the high target. 

2. Load one trial of your choice into R. The four columns refer to: sample, condition, x and y.

```{r}

knitr::opts_knit$set(root.dir = normalizePath("/Users/signeklovekjaer/Documents/CognitiveScience/4.semester/Computational_modeling_for_cognitive_science/ComMod/Models of perception and action/21"))


#load data from trial 1 
d1 <- read.delim("21_1.txt", sep = ",", header = FALSE)

#set names for columns
names(d1) <- c("sample", "condition", "x", "y")

```

3. Plot x against y. 

```{r}
#plot it 
ggplot(data = d1, aes(x,y)) + geom_point()
```

4. Map the y data to a more intuitive screen output with 0/0 in the lower left corner and where
upwards movements are increasingly positive (i.e. so that it looks like the movements we made
on the screen).


```{r}
#plot 
ggplot(data = d1, aes(x,y)) + geom_point() + ylim(0,550) + xlim(0,1700)

```

5. Calculate the distance travelled (for x and y combined).

```{r}
#calculate euclidian distance....
manualdist <- sum(sqrt(diff(d1$x)^2 + diff(d1$y)^2))
manualdist
```


6. Add a time vector. I forgot to measure time so I had to estimate it post-hoc (fine for this exercise,
not great if it was a real experiment). We can assume that each sample takes .0025 s. 

```{r}
2913*0.0025
time <- seq(from = 0, to = 7.28, by = 0.0025)

d1 <- cbind(d1, time)


```


7. Calculate velocity and acceleration for y, i.e. vertical movement. Remember: Velocity = distance
difference / time difference. Acceleration = velocity difference / time difference.

```{r}
#lag goes one back 

vel <- d1$y - lag(d1$y, k=1)
vel
```


```{r}
d1
```

