---
title: "hand_priming_exercises"
author: "Signe Kløve Kjær"
date: "3/4/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. 

Hypothesis: 

- In the condition with high targets and one low target, we expect the movement above the low target to be higher than the movement above the condition with only low targerts. 

- In the condition with low targets and one high targets, we expect the lower targets to be overshot relative to the condition with low targets --> preparing to move above the high target. 

2. Load one trial of your choice into R. The four columns refer to: sample, condition, x and y.

```{r}

knitr::opts_knit$set(root.dir = normalizePath("/Users/signeklovekjaer/Documents/CognitiveScience/4.semester/Computational_modeling_for_cognitive_science/ComMod/Models of perception and action/21"))


#load data from trial 1 
d1 <- read.delim("21_1.txt", sep = ",", header = FALSE)

#set names for columns
names(d1) <- c("sample", "condition", "x", "y")

```

3. Plot x against y. 

```{r}
library(ggplot2)
#plot it 
ggplot(data = d1, aes(x,y)) + geom_point()
```

4. Map the y data to a more intuitive screen output with 0/0 in the lower left corner and where
upwards movements are increasingly positive (i.e. so that it looks like the movements we made
on the screen).


```{r}
#plot 
ggplot(data = d1, aes(x,y)) + geom_point() + ylim(0,550) + xlim(0,1700)

```

5. Calculate the distance travelled (for x and y combined).

```{r}
#calculate euclidian distance....
manualdist <- sum(sqrt(diff(d1$x)^2 + diff(d1$y)^2))
manualdist
```


6. Add a time vector. I forgot to measure time so I had to estimate it post-hoc (fine for this exercise,
not great if it was a real experiment). We can assume that each sample takes .0025 s. 

```{r}
2913*0.0025
time <- seq(from = 0, to = 7.28, by = 0.0025)

d1 <- cbind(d1, time)


```


7. Calculate velocity and acceleration for y, i.e. vertical movement. Remember: Velocity = distance
difference / time difference. Acceleration = velocity difference / time difference.

```{r}
#lag goes one back 

vel <- (d1$y - dplyr::lag(d1$y, k=1)) / 0.0025
vel

d1 <- cbind(d1, vel)

acc <- (d1$vel - dplyr::lag(d1$vel, k=1)) / 0.0025

acc

d1 <- cbind(d1, acc)
```

8. Play around with some filter to smoothen velocity and acceleration (most common: Butterworth
filter).


```{r}
#change na to zero
d1[is.na(d1)] <- 0


#smooth velocity
lowpass.spline <- smooth.spline(d1$time,d1$vel, spar = 0.6) ## Control spar for amount of smoothing


#plot it 
plot(d1$time, d1$vel, type="l", lwd = 5, col = "green")
lines(predict(lowpass.spline, d1$time), col = "red", lwd = 2)
```
```{r}
#smooth acceleration

lowpass.spline_acc <- smooth.spline(d1$time,d1$acc, spar = 0.3) ## Control spar for amount of smoothing


#plot it 
plot(d1$time, d1$acc, type="l", lwd = 5, col = "green")
lines(predict(lowpass.spline_acc, d1$time), col = "red", lwd = 2)
```

9. Using zero crossings in the acceleration, calculate velocity extreme values (maxima and minima).
Mark those extreme values on top of the x/y plot you made earlier.
```{r}
vel_smooth <- predict(lowpass.spline, d1$time)

d1 <- cbind(d1, vel_smooth$y)

acc_smooth <- predict(lowpass.spline_acc, d1$time)

d1 <- cbind(d1, acc_smooth$y)

```


```{r}
#will be non-zero at cross-overs and will have the sign of the crossing
d1$zerocrossing <- c(0, diff(sign(d1$`acc_smooth$y`)))

ix <- which(d1$zerocrossing != 0)

d1$zerocrossing <- as.factor(ifelse(d1$zerocrossing == 0,0,1))

ggplot(data = d1, aes(x,y, color = zerocrossing,  alpha = zerocrossing)) + geom_point() + ylim(0,550) + xlim(0,1700)

```



```{r}

d1$zerocrossing_vel <- c(0, diff(sign(d1$`vel_smooth$y`)))


d1$zerocrossing_vel <- as.factor(ifelse(d1$zerocrossing_vel == 0,0,1))

ggplot(data = d1, aes(x,y, color = zerocrossing_vel, alpha = zerocrossing_vel)) + geom_point() + ylim(0,550) + xlim(0,1700)



```
```{r}
ix <- which(d1$zerocrossing_vel != 0)


table<-which(d1$zerocrossing_vel != 0)

d1$sixth<-ifelse(d1$sample< table[16] | d1$sample > table[18],0,1)

d1$sixth <-as.factor(d1$sixth)
ggplot(data=d1, aes(x=x,y=y,color=sixth)) +
  geom_point() +
  ylim(0,600) +
  xlim(0,1900)



```


QUESTION 9 FROM SIMONS CODE

```{r}
# 9.

ggplot(data=d1, aes(x=x,y=y,color=sixth, alpha = sixth)) + #plotting the data
  geom_point() +
  ylim(0,600) +
  xlim(0,1900) +
  scale_alpha_manual(guide='none', values = list('0' = 0.1, '1' = 0.1, '2' = 1))+
  scale_color_discrete(name = "", labels = c("Non-important trial", "Important trail", "y velocity changing direction"))

```


```{r}
# Bonus -->not changed to work for my data

# Here is a function that will run the analysis in question 10 for any set of data
mousePlot <-function(df){
  
  time <- seq(from=0,to=(nrow(df)-1)*0.0025,by=0.0025) #making a time seqence with the length of the dataframe with 0.0025 increments
  df<-cbind(df,time)
  
  colnames(df)<-c('sample','condition','x','y','time') #naming variables
  
  df$v<-((df$y-dplyr::lag(df$y, n= 1L))/0.0025) #calculating velocity
  
  df[is.na(df)] <- 0 
  
  lowpass.spline <- smooth.spline(df$time,df$v, spar = 0.3) #running a spline low pass filter
  sv<-predict(lowpass.spline, df$time) #running the filter on the data
  df<-cbind(sv$y,df) #saving the filtered data
  
  colnames(df)<-c('sv','sample','condition','x','y','time','v') #naming variables
  
  df$updn <- c(0, diff(sign(df$sv))) #finding were the sign switches velocity
  table<-which(df$updn!=0) #save the sample numbers of the sign switches

  for (n in (1:length(table))){ #finding the 6th movement
    if (df[table[n],'x']>1100 & df[table[n],'x']<1263)
      high<-n
  }
  
  df$imp<-ifelse(df$sample<table[high-1] | df$sample>table[high+1],0,1) #marking samples with the important 6th movement
  df$imp<-ifelse(df$updn==0,df$imp,2) #adding peaks to the 6th movement data to the dataframe
  df$imp <-as.factor(df$imp) #making it a factor
  
  ggplot(data=df, aes(x=x,y=y,color=imp, alpha = imp)) + #plotting the data
    geom_point() +
    ylim(0,600) +
    xlim(0,1900) +
    scale_alpha_manual(guide='none', values = list('0' = 0.1, '1' = 0.1, '2' = 1))+
    scale_color_discrete(name = "", labels = c("Non-important trial", "Important trail", "y velocity changing direction"))+
    labs(title=paste('Condition:' ,df[1,'condition']))
}

mousedata<-read.delim('21_2.txt',sep = ',', header = F) #loading data
mousePlot(mousedata) #running the function
```

